---
title: "GAGA LGTfinder analysis"
output:
  html_document:
    toc: true
    toc_float: true
    toc_depth: 2
    collapsed: true
    number_sections: true
    df_print: paged
  pdf_document: default
---
<STYLE TYPE="text/css">
<!--
  td{
    font-family: Arial; 
    font-size: 8pt;
    padding:0px;
    cellpadding="0";
    cellspacing="0"
  }
  th {
    font-family: Arial; 
    font-size: 8pt;
    height: 20px;
    font-weight: bold;
    text-align: right;
    background-color: #ccccff;
  }
  table { 
    border-spacing: 0px;
    border-collapse: collapse;
  }
--->
</STYLE>

# Load environment
```{r}
library(doParallel)
library(ggplot2)
library(GenomicRanges)
library(cowplot)
library(tidyverse)
library(ggbio)
library(raincloudplots)
library(ggrepel)
```

# Setup doParallel
```{r message=FALSE, warning=FALSE, include=FALSE}
cores<-detectCores()
cl <- makeCluster(cores[1]-1) #not to overload your computer
registerDoParallel(cl)

# necessary to load libraries again for %dopar% in all threads
clusterCall(cl, function() library(ggplot2))
clusterCall(cl, function() library(GenomicRanges))
clusterCall(cl, function() library(cowplot))
clusterCall(cl, function() library(tidyverse))
clusterCall(cl, function() library(ggbio))
clusterCall(cl, function() library(raincloudplots))
clusterCall(cl, function() library(ggrepel))
```

# load data
- lgt candidate loci
- lgt loci coverage info
- global coverage info
- swissprot hits in candidates

```{r}
GAGAid<-"GAGA-0515"
filepath<-paste("/Users/lukas/sciebo/Projects/LGT/results/",GAGAid,"/",sep="")
lgt.candidates<-read.csv(paste(filepath,"LGTs.candidateloci.bed",sep=""),sep="\t",F)
lgt.coverage<-read.csv(paste(filepath,"LGTs.candidateloci.coverage.bed",sep=""),sep="\t",F)
globalCoverage<-read.csv(paste(filepath,"genome.overlappingwindows.cov.tsv",sep=""),sep="\t",F)
lgt.proteins<-read.csv(paste(filepath,"LGTs.candidateloci.proteins.bed",sep=""),sep="\t",F)
#lgt.candidate.fasta<-"/Users/lukas/sciebo/Projects/LGT/results/GAGA-0515/LGTs.candidateloci.fa"
```

# Prepare data
## clean up lgt candidate loci information
**Colnames of the `lgt.candidates` dataframe are:**
 scaffold  start end bestProHit;tstart;tend;evalue;bits;alnlen;pident;taxname; collapsed(proBit)  collapsed(eukBit)  collapsed(bitDiff) collapsed(overlapping)

```{r}
colnames(lgt.candidates)<-c("cand.scaffold","cand.start","cand.end","bestProHit","cProBit","cEukBit","cBitDiff","cOverlap")
lgt.candidates$cand.locus<-paste(lgt.candidates$cand.scaffold,":",lgt.candidates$cand.start,"-",lgt.candidates$cand.end,sep="")
```

## clean up lgt loci coverage information
**Colnames of the `lgt.coverage` dataframe are:**
 scaffold  start end bestProHit;tstart;tend;evalue;bits;alnlen;pident;taxname;    coverage  basesCovered    windowsize
 
```{r}
colnames(lgt.coverage)<-c("scaffold","start","end","bestProHit","covWindowsStart","covWindowsEnd","cov","covBases","windowLength")
lgt.coverage$locus<-paste(lgt.coverage$scaffold,":",lgt.coverage$start,"-",lgt.coverage$end,sep="")
```

## clean up swissprot hit info
**Colnames of the `lgt.proteins` dataframe are:**
 scaffold  start end bestProHit;tstart;tend;evalue;bits;alnlen;pident;taxname; collapsed(proBit)  collapsed(eukBit)  collapsed(bitDiff) collapsed(overlapping)    hsp.scaffold   hsp.start hsp.end hsps(ProteinHit;tstart;tend;evalue;bits;alnlen;pident;taxname;, ...)    overlap.with.locus

```{r}
colnames(lgt.proteins)<-c("cand.scaffold","cand.start","cand.end","bestProHit","cProBit","cEukBit","cBitDiff","cOverlap","hsp.scaffold","hsp.start","hsp.end","hsps","overlap.w.locus")
lgt.proteins$cand.locus<-paste(lgt.proteins$cand.scaffold,":",lgt.proteins$cand.start,"-",lgt.proteins$cand.end,sep="")
lgt.proteins$hsp.locus<-paste(lgt.proteins$hsp.scaffold,":",lgt.proteins$hsp.start,"-",lgt.proteins$hsp.end,sep="")
```

## calculate genome wide median coverage
```{r}
colnames(globalCoverage)<-c("scaffold","windowstart","windowend","coverage","covBases","windowlength","fractionCovered")
averageGlobalCoverage<-median(globalCoverage$coverage)
```

## generate granges for overlapping 
```{r}
lgt.candidates.granges<-GRanges(seqnames=lgt.candidates$cand.locus)
lgt.coverage.granges<-GRanges(seqnames=lgt.coverage$locus)
overlaps<-as.data.frame(findOverlaps(lgt.coverage.granges,lgt.candidates.granges))
```

### Process data

Combine `lgt.candidates` with `lgt.coverage` into object `lgt`
```{r}
lgt<-cbind(lgt.candidates,lgt.coverage[overlaps$queryHits,])
```
 
Split `lgt` by candidates and `lgt.proteins` by hsp loci
```{r}
lgtL<-split(lgt,f = lgt$cand.locus)
lgt.proteinsL<-split(lgt.proteins,f = lgt.proteins$hsp.locus)
```
 
Create stacked data frame for  `lgt`
```{r}
lgt.stacked<-foreach(i=1:length(lgtL), .combine=rbind) %dopar% {
  tmp<-cbind(lgtL[[i]][c("cand.locus","scaffold","start","end","bestProHit","locus")],
             as.numeric(unlist(strsplit(x = lgtL[[i]]$covWindowsStart,split=","))),
             as.numeric(unlist(strsplit(x = lgtL[[i]]$covWindowsEnd,split=","))),
             as.numeric(unlist(strsplit(x = lgtL[[i]]$cov,split=","))),
             as.numeric(unlist(strsplit(x = lgtL[[i]]$covBases,split=","))),
             as.numeric(unlist(strsplit(x = lgtL[[i]]$windowLength,split=","))))
}
colnames(lgt.stacked)<-c("cand.locus","scaffold","start","end","bestProHit","locus","covWindowsStart","covWindowsEnd","cov","covBases","windowLength")
lgt.stacked$covWindowCenter<-(lgt.stacked$covWindowsEnd+lgt.stacked$covWindowsStart)/2

```

Create stacked data frame for  `lgt.protiens`
```{r}
# stack lgt.protein so each protein hit is one row
lgt.proteins.stacked.raw<-foreach(i=1:length(lgt.proteinsL), .combine=rbind) %dopar% {
  tmp<-cbind(lgt.proteinsL[[i]][c("cand.locus","cand.scaffold","cand.start","cand.end","bestProHit","hsp.start","hsp.end")],
            (unlist(strsplit(x = lgt.proteinsL[[i]]$hsps,split=","))))
}
colnames(lgt.proteins.stacked.raw)<-c("cand.locus","cand.scaffold","cand.start","cand.end","bestProHit","hsp.start","hsp.end","hsps")
lgt.proteins.stacked<-lgt.proteins.stacked.raw %>% separate(hsps, c("hit", "h.start","h.end","eval","bits","alnlen","pident","taxname"), fill = "right",sep = ";")
```

Split stacked `lgt` data frame by candidates

```{r}
lgt.stackedL<-split(lgt.stacked,f=lgt.stacked$cand.locus)
```
 
 
Loop over all candidate loci and plot

```{r}


#selection<-1
allPlots<-foreach(selection=1:length(lgt.stackedL)) %dopar% {

  lgt.candidates.subset<-subset(lgt.candidates,cand.locus == lgt.stackedL[[selection]]$cand.locus[1])
  lgt.proteins.subset<-subset(lgt.proteins.stacked,cand.locus == lgt.stackedL[[selection]]$cand.locus[1])
  lgt.proteins.subset.clean<-subset(lgt.proteins.subset,hit!=".")
  hsp.loci<-unique(lgt.proteins.subset[c("hsp.start","hsp.end")])
  unique.lgt.prot<-as.data.frame(lgt.proteins.subset.clean %>% group_by(hsp.start,hsp.end) %>% top_n(1, bits))
  BestBlastnHit<-unlist(strsplit(lgt.candidates.subset$bestProHit,";"))

    
  p1<-ggplot(lgt.stackedL[[selection]],aes(x=covWindowCenter,y=log(cov/averageGlobalCoverage,2)))+
    geom_point()+
    geom_line()+
    geom_smooth(se=F)+
    ylab("relCov")+
    theme_minimal_hgrid()+
    guides(x="none")+
    xlab(NULL)
    
  p2<-ggplot(lgt.stackedL[[selection]],aes(x=covWindowCenter,y=cov))+
    geom_bar(stat="identity",fill="grey80")+
    geom_segment(data=lgt.candidates.subset,aes(x=cand.start,xend=cand.end),y=0,yend=0,col="red",size=2,alpha=.9)+
    coord_cartesian(xlim=c(min(lgt.stackedL[[selection]]$covWindowCenter),max(lgt.stackedL[[selection]]$covWindowCenter)))+
    theme_minimal_hgrid()+
    xlab("NULL")
  

  blsPlot<-ggplot()+
          geom_segment(data=lgt.candidates.subset,aes(x=cand.start[1],xend=cand.end[1]),y=0,yend=0,col="red",size=2)+
          geom_text(data=lgt.candidates.subset,y=0.2,aes(x=cand.start[1]),label=paste(BestBlastnHit[8]," (",BestBlastnHit[1],")\ne-value = ",BestBlastnHit[4],sep=""),hjust=0)+
          geom_segment(data=unique.lgt.prot,aes(x=hsp.start,xend=hsp.end,color=as.numeric(bits)),y=.4,yend=.4,size=2)+
          geom_text_repel(data=unique.lgt.prot,aes(x=hsp.start,label=hit),y=.4,min.segment.length = unit(0, 'lines'),srt=90,size=4,box.padding = 1,direction="y",max.overlaps = Inf,nudge_y = .3, label.size =2,ylim=c(0.4,1.5))+
          coord_cartesian(ylim=c(0,1.5),xlim=c(lgt.candidates.subset$cand.start[1]-100,lgt.candidates.subset$cand.end[1]+100))+
          theme_minimal()+
          xlab(lgt.candidates.subset$cand.locus[1])+
          theme(legend.position=c(-0.02,.7),legend.title = element_text(size = 8))+
          guides(color = guide_colourbar(barwidth = .5, barheight = 4,title="bitscore"))


  # only continue if any protein hits are found
  if(unique(lgt.proteins.subset$hsp.start)!= "-1"){
    p3 <- ggplot(data=hsp.loci)+geom_rect(mapping=aes(xmin=hsp.start, xmax=hsp.end), ymin=0, ymax=1, color="red", alpha=0.1)+
      coord_cartesian(xlim=c(min(lgt.stackedL[[selection]]$covWindowCenter),max(lgt.stackedL[[selection]]$covWindowCenter)),ylim=c(0,1))+theme_nothing()
    
    
    lgt.prot.granges<-GRanges(seqnames = paste(lgt.proteins.subset.clean$cand.scaffold,":",lgt.proteins.subset.clean$hsp.start,"-",lgt.proteins.subset.clean$hsp.end,sep=""),
                              hit = lgt.proteins.subset.clean$hit,
                              bitscore=lgt.proteins.subset.clean$bits,
                              hit.start=lgt.proteins.subset.clean$h.start,
                              hit.end=lgt.proteins.subset.clean$h.end,
                              tax=lgt.proteins.subset.clean$taxname)
    
    lgt.prot.granges<-lgt.prot.granges[order(as.numeric(lgt.prot.granges$bitscore),decreasing = T)]
    
    alignmentPlot <- autoplot(lgt.prot.granges, aes(fill=as.numeric(bitscore),col=as.numeric(bitscore)),size=0)+
                      theme_classic()+theme(legend.position = c(-.02,.5),legend.title = element_text(size = 8))+
                      guides(fill = guide_colourbar(barwidth = .5, barheight = 4,title="bitscore"))+
                      coord_cartesian(xlim=c(min(lgt.stackedL[[selection]]$covWindowCenter),max(lgt.stackedL[[selection]]$covWindowCenter)))+
                      scale_color_continuous(guide = 'none')+
                      guides(y="none")
    
    alignmentPlotFocus<-autoplot(lgt.prot.granges[1,], fill="red2",size=0)+ylim(0,10)+annotate(geom="text", x=3, y=30, label="Scatter plot",
              color="red")
    
    
    tmpPlot<-plot_grid(blsPlot,p1,p2,alignmentPlot@ggplot,p3,NULL,ncol=1,align = "v",rel_heights = c(0.5,0.1,0.1,0.2,.01,0.01))
      
  }else{
      tmpPlot<-plot_grid(blsPlot,p1,p2,NULL,NULL,NULL,ncol=1,align = "v",rel_heights =c(0.3,0.3,0.3,0.02,0,0.01))
  }


}

```

```{r}
system(paste("mkdir ", filepath,"lgt.candidates",sep=""))
for (i in 1:length(allPlots)){
  ggsave(paste(filepath,"lgt.candidates/",GAGAid,".",gsub(":","-",names(lgt.stackedL)[i]),".pdf",sep=""),allPlots[[i]],width=6,height=10)
}
  

```

